Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2017-11-13 article.Rnw
Improve simulation study section

In this section we conducted a simulation study to assess the properties
of the proposed model and estimators. We considered average counts
varying from 3 to 27 according to a regression model structure with a
continuous variable $\bm{x}_1$ between 0 and 1 and a categorical
variable $\bm{x}_2$ with three levels. Then the systematic component is
given by $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
\bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where $\bm{x}_{21}$ and
$\bm{x}_{22}$ are \textit{dummy} variable for $\bm{x}_2$. With respect
to the precision parameter $\phi$, we fixed it to four values: $-1.6$,
$-1.0$, $0.0$ and $1.8$; in orderto have strong overdispersed, moderate
overdispersed, equidispersed and underdispersed counts
respectively. \autoref{fig:justpars} shows the variation of the average
counts (left) and variation of dispersion index (E$(Y)$/var$(Y)$) for
each fixed value of the precision parameter $\phi$. This configuration
allows us to evaluate the properties of the estimators in extreme
situations such as high counts and low dispersion, and low counts and
high dispersion.

In order to assess the consistency of the estimators we considered four
different sample sizes: 50, 100, 300 and 1000; generating 1000 datasets
in each case. In \autoref{fig:bias-plot} we show the bias of the
estimators for each scenario (combination between values of the
precision parameter and samples sizes) together with a confidence
interval calculated as average bias plus and minus $\Phi(0.975)$ times
the standard error. The scales are standardized for each parameter by
dividing the bias by the standard error obtained for the sample of size
50.

@@

@@

The results in \autoref{fig:bias-plot} show that for all dispersed
levels, both the expected bias and standard error tend to 0 as the
sample size is increased. The estimators for the regression parameters
are unbiased, consistency and their empirical distributions are
symmetric. For the precision parameter, the estimator is asymptotically
unbiased, in small samples the parameter is overestimated and the
empirical distribution is slightly right-skewed.
\autoref{fig:coverage-plot} presents the confidence interval coverage
rate based on quadratic approximation at the maximum likelihood estimate
(assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
\sqrt{\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
observed information matrix) by sample size and dispersed levels.

@@

The coverage rates in \autoref{fig:coverage-plot} show that for the
regression parameters the empirical coverage rates are close to the
nominal level of 95\% for sample sizes greater than 100 and all
dispersed levels. For the precision parameter the empirical coverage
rates are slightly lower than the nominal level, however, they become
closer for large samples. The worst scenario is when we have small
sample size and strong overdispersed counts.

>>>>

In this section we conducted a simulation study to assess the properties
of the maximum likelihood estimators as well as the flexibility of the
COM-Poisson regression model to deal with non-equidispersed count data.

We considered average counts varying from $3$ to $27$ according to a
regression model with a continuous and a categorical covariate. The
continuous covariate~($\bm{x}_1$) was generated as a sequence from $0$
to $1$ and length equals to the sample size.  Similarly, the categorical
covariate~($\bm{x}_2$) was generated as a sequence of three values each
one repeated $3/n$ times (rounding up when required), where $n$ denotes
the sample size. Thus, the expectation of the COM-Poisson random
variable is given by
$\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21} \bm{x}_{21} +
\beta_{22} \bm{x}_{22})$,
where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels
of $\bm{x}_2$.  The regression coefficients were fixed at the values,
$\beta_0 = 2$, $\beta_1 = 05$, $\beta_{21} = 0.8$ and
$\beta_{22} = -0.8$.

We designed four simulation scenarios by considering four different
values of the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.
Thus, we have strong and moderate overdispersion, equidispersion and
underdispersion, respectively.  \autoref{fig:justpars} shows the
variation of the average counts (left) and dispersion index (right) for
each value of the dispersion parameter considered in the simulation
study.  These configurations allow us to assess the properties of the
maximum likelihood estimators in extreme situations such as high counts
and low dispersion, and low counts and high dispersion, but also in the
standard case of equidispersion.

@@

In order to check the consistency of the estimators we considered four
different sample sizes: $50$, $100$, $300$ and $1000$; generating $1000$
data sets in each case.  In \autoref{fig:bias-plot} we show the bias of
the estimators for each simulation scenario (combination between values
of the dispersion parameter and samples sizes) along with the confidence
intervals calculated as average bias plus and minus $\Phi(0.975)$ times
the average standard error. The scales are standardized for each
parameter by dividing the average bias by the average standard error
obtained for the sample of size $50$.

@@

The results in \autoref{fig:bias-plot} show that for all dispersion
levels, both the average bias and standard errors tend to $0$ as the
sample size is increased. The estimators for the regression parameters
are unbiased, consistency and their empirical distributions are
symmetric. For the dispersion parameter, the estimator is asymptotically
unbiased, in small samples the parameter is overestimated and the
empirical distribution is slightly right-skewed.
\autoref{fig:coverage-plot} presents the empirical coverage rate of the
asymptotic confidence intervals.

@@

The coverage rates in \autoref{fig:coverage-plot} show that for the
regression parameters the empirical coverage rates are close to the
nominal level of 95\% for sample sizes greater than $100$ and all
simulation scenarios. For the dispersion parameter the empirical
coverage rates are slightly lower than the nominal level, however, they
become closer to the nominal level for large samples.  The worst
scenario is when we have small sample size and strong overdispersed
counts.
